\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-08-26 gwnnDraft02.tex
more minor edits
new plots
email address
added additional refs and fixed some minor points

\email{
PaulSmith3@cmail.carleton.ca \\
}

@@

Because gravity is universal, gravitational waves can allow us to probe hidden sectors that couple very weakly, or not at all, to the Standard Model as long they are reheated after inflation. This was first explored in~\cite{Schwaller:2015tja}, and there has been significant work on this idea since~\cite{Breitbach:2018ddu,}.

@@

The SM predicts that $N^{SM}_{eff} = 3.046$ \cite{Mangano:2005cc}. This is in good agreement with the $2\sigma$ bounds from studies of the Cosmic Microwave Background (CMB) by Planck combined with baryon acoustic oscillations (BAO) measurements \cite{Aghanim:2018eyx}:
\begin{equation}\label{eqn:NeffBounds}
N_{eff} = 2.99^{+0.34}_{-0.33}.
\end{equation}
Various different assumptions about the history of the universe can be made and different data sets can be chosen to obtain slightly different results \cite{Breitbach:2018ddu} --- for the purposes of this exploratory work, wading through this landscape is unnecessary. Additionally, 
\begin{equation}
\frac{(\Delta N^i_{eff})_{CMB}}{(\Delta N^i_{eff})_{BBN}} \geq 1
\end{equation}
for any \draft{check} decoupled hidden sector~\cite{Arkani-Hamed:2016rle}. Because the constraints on $N_{eff}$ are stronger at photon decoupling than at BBN, we can focus purely on the constraints provided by the former.

Future CMB experiments~\cite{} will improve the bound from Eq.~(\label{eqn:NeffBounds}) by about an order of magnitude. This could significantly reduce the allowed temperature ratio of any hidden sector, or alternatively could provide evidence for such sectors in a way that is complementary to the gravitational wave signatures described below.

@@

We begin by computing the constraints on exotic sectors; these are significantly weaker than those for standard sectors~\cite{Arkani-Hamed:2016rle}. 
%At high scales, these sectors feature 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. 
At the time of photon decoupling ($T_\gamma \sim 0.39 \,\ev$) while the temperature of the exotic sectors is lower still. This means that for sectors with small and moderate $i$, we can use Eqs.~(\ref{eqn:exoticFM}) and~(\ref{eq:ith_pion_mass}) to see that the pions will be non-relativistic  leaving at most $11.25$ effective degrees of freedom per sector from photons and neutrinos. \draftnote{Pions are not decoupled at BBN time, does this in any way affect the analysis?} For very large $i$, the pions can be much lighter, but those sectors also have very little energy in them in the standard reheating scenario.

@@

\draftnote{What is temperature ratio for random reheating scenario?}

@@

Using these restrictions, we can examine the behaviour of standard sectors with a much larger vev than our own. In terms of the $N$naturalness framework, this means we can get an SFOPT for QCD if we look at sectors with $i$ greater than the cirtical index of Eq.~(\ref{eqn:critIndex}) where all the quark masses are above the QCD confinement scale. 
 %of looking at various different sectors above the critical index, as mentioned in Sec.~\ref{sec:dQCD}. The signals for both the maximal case ($10^{16}$ limit) and several other high transition temperatures (involving sectors $i > 10^{10}$) scenarios are shown in Sec.~\ref{sec:gw}. \textcolor{yellow}{This assumes that only photons and neutrinos are relativistic? Yes.}

>>>>

\email{
PaulSmith3@cmail.carleton.ca \\
Dylan.linthorne@carleton.ca \\
stolar@physics.carleton.ca
}

@@

Because gravity is universal, gravitational waves can allow us to probe hidden sectors that couple very weakly, or not at all, to the Standard Model as long they are reheated after inflation. This was first explored in~\cite{Schwaller:2015tja}, and there has been significant work on this idea since~\cite{Breitbach:2018ddu,Bertone:2019irm,Caputo:2019wsd,Helmboldt:2019pan,Fairbairn:2019xog,Bai:2018dxf,Baldes:2018emh,Croon:2018erz,Geller:2018mwu,Heikinheimo:2018esa,Aoki:2017aws,Bernal:2017kxu,Acharya:2017szw,Tsumura:2017knk,Dienes:2016vei,Hardy:2016mns,Addazi:2016fbj,Jaeckel:2016jlh}.

@@

The SM predicts that $N^{SM}_{eff} = 3.046$ \cite{Mangano:2005cc}. This is in good agreement with the $2\sigma$ bounds from studies of the Cosmic Microwave Background (CMB) by Planck combined with baryon acoustic oscillations (BAO) measurements \cite{Aghanim:2018eyx}:
\begin{equation}\label{eqn:neffBounds}
N_{eff} = 2.99^{+0.34}_{-0.33}.
\end{equation}
Various different assumptions about the history of the universe can be made and different data sets can be chosen to obtain slightly different results \cite{Breitbach:2018ddu} --- for the purposes of this exploratory work, wading through this landscape is unnecessary. Additionally, 
\begin{equation}
\frac{(\Delta N^i_{eff})_{CMB}}{(\Delta N^i_{eff})_{BBN}} \geq 1
\end{equation}
for any \draft{check} decoupled hidden sector~\cite{Arkani-Hamed:2016rle}. Because the constraints on $N_{eff}$ are stronger at photon decoupling than at BBN, we can focus purely on the constraints provided by the former.

Future CMB experiments~\cite{Abazajian:2016yjj} will improve the bound from Eq.~(\ref{eqn:neffBounds}) by about an order of magnitude. This could significantly reduce the allowed temperature ratio of any hidden sector, or alternatively could provide evidence for such sectors in a way that is complementary to the gravitational wave signatures described below.

@@

We begin by computing the constraints on exotic sectors; these are significantly weaker than those for standard sectors~\cite{Arkani-Hamed:2016rle}. 
%At high scales, these sectors feature 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. 
At the time of photon decoupling ($T_\gamma \sim 0.39 \,\ev$) while the temperature of the exotic sectors is lower still. This means that for sectors with small and moderate $i$, we can use Eqs.~(\ref{eqn:exoticFM}) and~(\ref{eq:ith_pion_mass}) to see that the pions will be non-relativistic  leaving at most $11.25$ effective degrees of freedom per sector from photons and neutrinos.  For very large $i$, the pions can be much lighter, but those sectors also have very little energy in them in the standard reheating scenario.

@@

@@

Using these restrictions, we can examine the behaviour of standard sectors with a much larger vev than our own. In terms of the $N$naturalness framework, this means we can get an SFOPT for QCD if we look at sectors with $i$ greater than the cirtical index of Eq.~(\ref{eqn:critIndex}) where all the quark masses are above the QCD confinement scale.

 %of looking at various different sectors above the critical index, as mentioned in Sec.~\ref{sec:dQCD}. The signals for both the maximal case ($10^{16}$ limit) and several other high transition temperatures (involving sectors $i > 10^{10}$) scenarios are shown in Sec.~\ref{sec:gw}. \textcolor{yellow}{This assumes that only photons and neutrinos are relativistic? Yes.}
